---
title: "Capstone Adult Data"
author: "Mark Stallworth"
date: "February 5, 2018"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Business Understanding: Goal is to predict income greater than 50K based on attributes of the adult.data.txt dataset

# Load the required packages and libraries
```{r message=FALSE, echo=FALSE}
#install.packages('caret', dependencies = TRUE)
#install.packages('readr')
#install.packages('dplyr')
#install.packages('e1071')
#install.packages('mice')
#install.packages('gdata')
#install.packages('psych')
#install.packages('DMwR',dependencies = TRUE)
library(caret)
library(readr)
library(e1071)
library(ggplot2)
library(dplyr)
library(DMwR)
library(mice)
library(gdata)
library(psych)
```

2. Data Understanding
```{r message=FALSE, echo=FALSE}
#Read the adult dataset into data.frame rawData without lables
rawData <- read.table("C:/Springboard/DS Fundamentals/R Code/Capstone/data/adult.data.txt", header = FALSE, sep = ",")

#Assign Labels
names(rawData) <- c("age", "workclass", "fnlwwgt", "education", "education_num", "marital_status", "occupation", "relationship", "race", "sex", "capital_gain", "capital_loss", "hours_per_week", "native_country", "income50")
```

#Review data structures and browse data
```{r message=FALSE, echo=FALSE}
str(rawData)
class(rawData)
View(rawData)
```

# Data Exploration Factor Variables
```{r message=FALSE, echo=FALSE}  
workclass_plot <- ggplot(rawData, aes(workclass), stat_count()) + geom_bar()
workclass_plot
education_plot <- ggplot(rawData, aes(education), stat_count()) + geom_bar()
education_plot
marital_status_plot <- ggplot(rawData, aes(marital_status), stat_count()) + geom_bar()
marital_status_plot
occupation_plot <- ggplot(rawData, aes(occupation), stat_count()) + geom_bar()
occupation_plot
relationship_plot <- ggplot(rawData, aes(relationship), stat_count()) + geom_bar()
relationship_plot
race_plot <- ggplot(rawData, aes(race), stat_count()) + geom_bar()
race_plot
sex_plot <- ggplot(rawData, aes(sex), stat_count()) + geom_bar()
sex_plot
native_country_plot <- ggplot(rawData, aes(native_country), stat_count()) + geom_bar()
native_country_plot
```
# Data Exploration Factor Variables with "?" in columns of observations select counts
```{r message=FALSE, echo=FALSE}
select(rawData) %>% group_by(rawData$workclass) %>% summarise(count = n()) 
select(rawData) %>% group_by(rawData$occupation) %>% summarise(count = n())
select(rawData) %>% group_by(rawData$native_country) %>% summarise(count = n())
```

# Check for missing values in integer columns:
```{r message=FALSE}
sum(is.null(rawData$age))
sum(is.null(rawData$fnlwwgt))
sum(is.null(rawData$education_num))
sum(is.null(rawData$capital_gain))
sum(is.null(rawData$capital_loss))
sum(is.null(rawData$hours_per_week))
```
3 Data Preperation
```{r message=FALSE}
#identify factor variables with potential missing data and exclude label income50
factor_idx = which(unlist(lapply(rawData, class)) == "factor")
factor_idx_depvar <- (factor_idx[-9])  #remove column income50

#Identify relative column number for factors to be used to replace "?" with "NA"
for (i in 1:NROW(factor_idx_depvar)) {
  naidx <- which(as.character(rawData[, factor_idx_depvar[i]]) == " ?")
  rawData[naidx, factor_idx_depvar[i]] <- NA
}
#Drop category levels in order to remove "?" level and replace with NA; other wise imputaion substitutes "?" from original data
rawData <- drop.levels(rawData)

#replace NA values with K nearest neighbor
?knnImputation
rawDataClean <- knnImputation(rawData, 4)

#Verify clean data for the 3 columns (workclass, occupation and native_country) that contains missing data in source input file (adult.data.txt)
select(rawDataClean) %>% group_by(rawDataClean$workclass) %>% summarise(count = n()) 
select(rawDataClean) %>% group_by(rawDataClean$occupation) %>% summarise(count = n())
select(rawDataClean) %>% group_by(rawDataClean$native_country) %>% summarise(count = n())
```
4. Modeling
```{r message=FALSE, echo=FALSE}
preProcModel = preProcess(rawDataClean, method = c("center", "scale", "zv", "nzv"))
processedData = predict(preProcModel, rawDataClean)
class(processedData)
dim(processedData)

#Create dummy codes for factors, this is required for svm modeling
processedData$workclass <- dummy.code(processedData$workclass)
processedData$education <- dummy.code(processedData$education)
processedData$marital_status <- dummy.code(processedData$marital_status)
processedData$occupation <- dummy.code(processedData$occupation)
processedData$relationship <- dummy.code(processedData$relationship)
processedData$race <- dummy.code(processedData$race)
processedData$sex <- dummy.code(processedData$sex)
processedData$native_country <- dummy.code(processedData$native_country)
```
```{r message=FALSE}
trainRows = sample(seq(1,nrow(processedData)),0.75*nrow(processedData))
length(trainRows)

trainData = processedData[trainRows,]
dim(trainData)

#testRows = setdiff(seq(1,nrow(inputData), trainRows)
testRows = setdiff(seq(1, nrow(processedData)), trainRows)

testData = processedData[testRows,]
dim(testData)

#Process a sample of training rows due to processing constraings and educational objective per Mentor
trainData <- trainData[sample(1:nrow(trainData), 5000, replace=FALSE),]
nrow(trainData)

initialModel = svm(x=trainData[,1:(ncol(trainData)-1)],y=trainData[,ncol(trainData)])
summary(initialModel)

initialModelPredict <- predict(initialModel,trainData[,1:(ncol(trainData)-1)])
str(initialModelPredict)
NROW(initialModelPredict)

svm.table <- table(initialModelPredict, trainData$income50)
classAgreement(svm.table)
confusionMatrix(svm.table)

tunedInitialModel2 <- tune(svm, train.x = trainData[,-13], train.y=trainData$income50, kernel="radial", ranges=list(cost=exp(c(-1,0,3,5,7,9)), gamma=exp(c(-1,0,3,5,7,9))))
summary(tunedInitialModel2)
print(tunedInitialModel2)
```













